---
title: "LS6_DataPresentation"
authors: "Kelly Bean, Olivia Leeson, John Dorsey"
date: "October 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
########################
#
# This portion of R code creates
# objects used in later analyses.
#
########################
library(plyr)
library(gdata)
setwd("C:\\Users\\Kelly\\Desktop\\ClassFall2016\\Doing Data Science")

#Save csv data to single object 'bk'
bk <- read.csv("rollingsales_statenisland2.csv",skip=4,header=TRUE)
## Create variable sale.price.n as numeric
## Use the gsub command to replace digits with a blank space
bk$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", bk$SALE.PRICE))
#count(is.na(bk$SALE.PRICE.N))

## Make all variable names lower case
names(bk) <- tolower(names(bk))

# Create new variables:
# Get rid of leading digits
# Make factors numeric values
bk$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$gross.square.feet))
bk$land.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$land.square.feet))
bk$year.built <- as.numeric(as.character(bk$year.built))

## New dataset for data on properties that have sales
bk.sale <- bk[bk$sale.price.n!=0,]

## New dataset for 1,2,3 bedroom family homes
bk.homes <- bk.sale[which(grepl("FAMILY",bk.sale$building.class.category)),]

## Remove outliers in sales data
bk.homes$outliers <- (log10(bk.homes$sale.price.n) <=5) + 0
bk.homes <- bk.homes[which(bk.homes$outliers==0),]

## Format dates
bk.homes$saledate <- as.character(bk.homes$sale.date)
bk.homes$saledate <- as.Date(bk.homes$saledate,"%m/%d/%y")
```


# Meaningful Patterns
The most obvious pattern presented by the data set is the positive relationship between square footage and sale price; as square footage increases, so does the sale price.
Even performing a log-transformation on the data to count for right-skewness, the data still show a generally positive relationship. 

The scatterplots below illustrate this relationship; the first scatterplot is on original data for all homes, the second is on log-transformed data only for 1-, 2-, and 3-family homes.
```{r}
plot((bk.homes$gross.sqft),(bk.homes$sale.price.n), 
     main="Square Footage vs. Sale Price", xlab='Square Footage', ylab='Sale Price')

bk.homes <- bk.sale[which(grepl("FAMILY",bk.sale$building.class.category)),]
plot(log10(bk.homes$gross.sqft),log10(bk.homes$sale.price.n), 
     main="Log Transformed Sqft vs. Sale Price", xlab='Log10(SqFt)', ylab='Log10(SalePrice)')
```

